\documentclass[11pt]{article}
\usepackage[a4paper,margin=1in]{geometry}
\usepackage{lmodern}
\usepackage{titlesec}
\usepackage{enumitem}
\usepackage{hyperref}
\usepackage{array}
\usepackage{xcolor}
\usepackage{cleveref}
\usepackage{graphicx}

\usepackage{tikz}
\usetikzlibrary{positioning, shapes, arrows.meta}

\titleformat{\section}{\large\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\normalsize\bfseries}{\thesubsection}{1em}{}

\title{\textbf{Oxbridge Maths Preparation Guide}}
\date{\today}


% --- Options ---
\newif\ifshowcomments
\newif\ifshowpersonal
\newcommand{\studentname}{Anonymous}
\showcommentstrue
\showpersonaltrue
\renewcommand{\studentname}{Aarna}
% --------------

\usepackage{xcolor}

% --- Custom Colors ---
\definecolor{myblue}{RGB}{30, 90, 180}
\definecolor{myorange}{RGB}{230, 120, 30}

\newcommand{\comment}[1]{\ifshowcomments\textcolor{myblue}{\textbf{#1}}\fi}
\newcommand{\personal}[1]{\ifshowpersonal\textcolor{myorange}{\emph{#1}}\fi}


\begin{document}

\maketitle

\bigskip

%--------------------------------------------------
\section{Introduction}
\noindent
The purpose of what follows is threefold. Firstly it describes the unique features of Oxbridge maths, to help you decide how much you want to invest in an application to Oxbridge. Secondly it then lays out the what the application actually involves. Finally it provides resources and tips and tricks to ensure that your application is as competitive as it can be!\footnote{This is largely an amalgamation of resources from various sources and most are not my own!!}  

\ifshowcomments
To make it a bit more personable I've added comments specific to you \comment{(\studentname)} in \comment{bold}.
\fi
\ifshowpersonal 
Comments on what my own experience was like are in \personal{italics}.
\fi


%--------------------------------------------------
\section{Why Oxbridge?}

Oxbridge suits a wide range of people, and there are many reasons why it’s such a special place to study. The academic excellence and the unique tutorial or supervision system are certainly key draws, as is the world-class research and faculty. The collegiate environment and traditions create a sense of community and continuity that few other universities can match. And then, of course, there’s the challenge. It’s not an easy place to study — but for some, that’s exactly the point.

Put simply, there’s nowhere else quite like it. But I should also say that there are perfectly good reasons not to apply to Oxford. The brightest students tend to go there, so it’s hardly surprising that Oxford produces clever graduates — but I wouldn’t necessarily credit the teaching alone for that. These are people who would likely succeed anywhere. There are brilliant minds all around, some of whom will be teaching you, and if you seek them out, you can learn a lot. But only a small fraction of students — maybe ten percent — actually take advantage of that.

\personal{Personally, I probably didn't. However I did always try to read widely around the subject.}

Perhaps the top thirty percent of students end up thriving academically at Oxford. For many others, it can be a demoralising experience — not because they aren’t capable, but because the environment doesn’t always instil enthusiasm. That’s not to say they don’t enjoy the university itself — it’s a wonderful place to be a student — but many people I know leave with a hate of studying their subject. This is obviously not conducive to them doing well at it.

Doing well at Oxford doesn’t necessarily require more work than elsewhere, but the culture can make it feel that way. People are older than they were in school, and what used to be a normal workload might now feel excessive, just because everyone is approaching it differently. It’s quite a shift. This isn’t universal — on the continent, for instance, universities can be much more competitive in structure. Entry is often open, but in many maths courses, up to sixty percent of students are removed after the first semester.

So I’d say it’s entirely possible to have a more enjoyable, maybe even more fulfilling, experience at another university — and to come out of it with a stronger love of maths and, arguably, similar job prospects. That said, unless you do exceptionally well elsewhere, you probably won’t learn as much maths in total. Then again, you may well find it easier to learn in that environment. The pressure and academic intensity at Oxford are very real, and they affect people in different ways.

\comment{From what I've seen of you, Jack, you absolutely have the curiosity and ability that Oxbridge looks for. I wouldn’t worry about whether you’re the right fit — you are. It’ll just be a matter of keeping a focus on your studies while you're there.}

\section{The application process}

The Oxbridge admissions process works remarkably well. Very few people are admitted who don’t belong there — but it’s entirely possible for someone who should get in to be rejected. Because of that, and because of the academic intensity once you arrive, I would encourage you not just to aim to scrape in, but to be in the group of applicants for whom rejection is essentially impossible. The closer you are to that level, the more prepared you’ll be to thrive once you’re there.

Getting to that point requires serious work. If you decide that getting into Oxbridge is something truly important to you, then — concretely — it means making it a major focus of your life over the coming months. In practice, I’d say that aside from mealtimes, daily exercise, and good-quality sleep (which is crucial — consistency of bedtime matters even more than duration, though eight hours is a good target), you shouldn’t be giving yourself more than two hours of downtime per day. This is a key period, and one where I personally found it helpful to cut down on the usual things that eat time for little reward.

If you take a 24-hour day and subtract eight hours for sleep, two for meals and life admin, six for school, and another two for downtime, that still leaves you with six hours a day to dedicate to maths. That’s a lot — but not unmanageable.

Now, that doesn’t mean six hours of head-in-hands, intense problem-solving every day. A lot of it should be focused and challenging, but there are plenty of ways to bring variety into your study routine — some of which I’ll outline later. The key is to stay consistent and keep your mind immersed in mathematical thinking.

\personal{For me, this meant cutting out things like watching TV shows. That was the approach I chose. I'm not claiming it’s what’s needed to get into Oxford — I almost certainly could have done less and still made it — but it proved very useful during my time at Oxford, and I enjoyed it.}

\comment{That said, if the idea of that routine fills you with dread — and you can’t imagine sustaining it even for a month — then I’d take a moment to think about how much you’d actually enjoy Oxford. You don’t have to love the idea of that much maths every day, and the workload at university may well be less structured or even less intense at times. But it’s worth asking: do you enjoy the subject enough to want to live and breathe it for a while? If the answer is yes, you’ll probably be just fine.}

Universities build a picture of you as a student from all the different information you provide, to help decide whether or not to offer you a place.

The picture is made up of several different pieces:

    your personal statement,
    your academic record (this is contextualised based on the school you attended),
    your predicted A-level grades (or equivalent),
    your entrance exam (either MAT or STEP),
    and your teacher's reference.

 

%--------------------------------------------------
%--------------------------------------------------
\subsection{Timeline of Application Process}
\begin{center}
\scalebox{0.85}{
\begin{tikzpicture}[>=Stealth]
\usetikzlibrary{calc}
% --- Top row (left to right) ---
\coordinate (prep)      at (0,0);
\coordinate (ucasopen)  at (4,0);
\coordinate (clarus)    at (8,0);
\coordinate (booking)   at (12,0);

% --- Downward turn ---
\coordinate (turn)      at (15,-2.5);

% --- Bottom row (right to left) ---
\coordinate (ucasclose) at (12,-5);
\coordinate (matdates)  at (8,-5);
\coordinate (interviews)at (4,-5);
\coordinate (offers)    at (0,-5);

% --- Draw arrows ---
\draw[thick] (prep) -- (booking);              % Top row
\draw[thick] (booking) to[out=0, in=90] (turn) to[out=270,in=0] (ucasclose);
\draw[thick] (ucasclose) -- (offers);               % Bottom row

% --- Tick marks ---
\foreach \pt in {prep, ucasopen, clarus, booking, ucasclose, matdates, interviews, offers}
  \draw[thick] (\pt) -- ++(0,0.3);

% --- Dates above ---
\node[align=center, anchor=south] at ($(prep)+(0,0.3)$) {\textbf{Now–Aug}};
\node[align=center, anchor=south] at ($(ucasopen)+(0,0.3)$) {\textbf{Early Sep}};
\node[align=center, anchor=south] at ($(clarus)+(0,0.3)$) {\textbf{18 Jun–19 Sep}};
\node[align=center, anchor=south] at ($(booking)+(0,0.3)$) {\textbf{18 Aug–26 Sep}};

\node[align=center, anchor=north] at ($(ucasclose)-(0,0.3)$) {\textbf{15 Oct, 6pm}};
\node[align=center, anchor=north] at ($(matdates)-(0,0.3)$) {\textbf{22–23 Oct}};
\node[align=center, anchor=north] at ($(interviews)-(0,0.3)$) {\textbf{Dec}};
\node[align=center, anchor=north] at ($(offers)-(0,0.3)$) {\textbf{Jan 2026}};

% --- Descriptions below/above ---
\node[align=center, anchor=north, text width=4cm] at ($(prep)-(0,0.2)$) {
Begin prep: reading,\\MAT/interview practice,\\personal statement
};
\node[align=center, anchor=north, text width=4cm] at ($(ucasopen)-(0,0.2)$) {
UCAS portal opens:\\start application
};
\node[align=center, anchor=north, text width=4cm] at ($(clarus)-(0,0.2)$) {
Register for tests on\\\href{https://oxford.useclarus.com}{oxford.useclarus.com}
};
\node[align=center, anchor=north, text width=4cm] at ($(booking)-(0,0.2)$) {
Book test centre\\via Clarus
};

\node[align=center, anchor=south, text width=4cm] at ($(ucasclose)+(0,0.2)$) {
	Submit \href{https://accounts.ucas.com/account/login}{UCAS}\\application
};
\node[align=center, anchor=south, text width=4cm] at ($(matdates)+(0,0.2)$) {
Sit the MAT on 22–23 Oct\\\href{https://www.ox.ac.uk/admissions/undergraduate/applying-to-oxford/guide/admissions-tests/mat}{ox.ac.uk/mat}
};
\node[align=center, anchor=south, text width=4cm] at ($(interviews)+(0,0.2)$) {
Interviews at colleges\\(if shortlisted)
};
\node[align=center, anchor=south, text width=4cm] at ($(offers)+(0,0.2)$) {
Offers sent out\\ Jan 2026
};

\end{tikzpicture}
}
\end{center}
%--------------------------------------------------

\subsection{UCAS application}

There are things you can change and some you can't.

First, for Maths at universities like Cambridge, Warwick, and Oxford (most high-ranking ones but not all),
they won’t really read or care much about your personal statement. For example, Warwick claims they
briefly look at it for a few minutes and never look at it again, and I’ve heard Trinity have said that they
read it just because they have to. This is because they have better means to measure the suitability of a
candidate for their course, such as interviews and admissions tests.
What you should take away from this isn’t that you don’t need to write a good personal statement but that
you don’t need to worry about doing extremely specific extracurriculars or very minute details. Make sure
the statement is good for other universities. James Munro, the admissions director at Oxford, put it best
when he said, “If you have an hour to either improve your personal statement or do some maths, you should
do some maths.”
They’re generally not interested at all in anything in your personal statement which isn’t about maths, but
it is good to have a line or two on this for the sake of other universities. In the interview, they’re generally
unlikely to ask you about anything on your personal statement, but it’s worth being aware of what you wrote
in case they ask you about something related to it. If you claim you know a lot of university maths, you
should know it! You should also know a little bit around it too. To take an example from university maths:
if you say you know Cauchy’s integral theorems, then you should know Liouville’s theorem, an important
(but quite quick) deduction from Cauchy’s integral formula.
However, this doesn’t mean that doing extracurricular maths is a bad thing. In fact, it’s probably the best
thing you can be doing at this stage – it’s just that you shouldn’t do things just for the personal statement.
Instead, explore the maths you enjoy the most and want to learn about. That way, you’ll learn a lot more
and end up organically having a lot of things to write on your personal statement later on, which will make
writing it a lot easier.

\href{https://www.undergraduate.study.cam.ac.uk/apply/how/ucas-personal-statement}{Advice from Cambridge} on personal statements. 

\href{https://www.ox.ac.uk/admissions/undergraduate/applying-to-oxford/guide/ucas-application}{Advice from Oxford} on personal statements.

\subsection{Written exams}


\subsection{Interviews}



\href{https://www.ox.ac.uk/admissions/undergraduate/applying-to-oxford/guide/interviews}{Interviews}



\section{Resources}

This is by no means a complete list — rather, it’s a collection of things I did which I found interesting or useful in some way. If you'd like more detail on any of the sections, feel free to ask. The order is not significant — in fact, you might find some of the later items the most enjoyable or valuable.

\subsection{Targeted MAT Practice}

\begin{enumerate}
\item \href{https://www.maths.ox.ac.uk/study-here/undergraduate-study/maths-admissions-test}{MAT past papers}. While the syllabus and format have changed recently, the old papers remain an excellent source of problems and give you the clearest idea of what to expect. I’d recommend starting with the earlier ones and working your way forward.

\item \href{https://www.colmanweb.co.uk/Assets/PDF/advanced-problems-mathematics.pdf}{Stephen Siklos' STEP Book}. STEP past paper questions are also brilliant practice, though they can be very challenging at first. That’s why this book is helpful — it includes worked solutions, and helps bridge the gap into STEP-style thinking.
\end{enumerate}

\subsection{Interview Practice}

This was one of the most confusing parts of the application process for me, as it’s hard to know exactly what to expect. I’d strongly encourage you to ask teachers or mentors for mock interviews, and to practise talking through your reasoning aloud. It is also very useful to go on youtube and search for recordings of past interviews, for example \href{}{}.

\begin{itemize}

\item \href{https://www.drfrost.org/uploads/users/1/r261/TBO\%20Problem\%20Solving\%20Booklet.pdf}{\textbf{TBO Problem Solving Booklet}} --- some very good problems in the style of interview questions. 

\item \herf{https://vg354.user.srcf.net/resources/Interview_questions.pdf}{\textbf{Example interview questions}} --- some more past interview questions.

\item \herf{https://www.cl.cam.ac.uk/~abr28/prob/questions.pdf}{\textbf{Probability interview questions}} --- by the creator of the CSAT, some hard probability interview style questions.

\end{itemize}

\subsection{Books}

These are a selection of more technical books, wider reading is also encouraged in an effort to built a broader interest around the subject - suggests for this are included in \cref{sec:morebooks}.

\begin{itemize}
\item \textbf{What Is Mathematics?} by Richard Courant and Herbert Robbins — a classic text that presents “older” mathematical ideas in a very elegant way. \emph{Let me know if you’d like a PDF copy.}

\item \href{https://ukmt.org.uk/product/a-problem-solvers-handbook}{\textbf{UKMT Problem Solver’s Handbook}} — one of a wider series of books from UKMT. This contains clear explanations and rich problems. \emph{I think I may have even started with the junior-level book (ages 11–13) when I was 16, since I hadn’t encountered much of this style before.}

\item \href{https://isaacphysics.org/books/pre_uni_maths_2e}{\textbf{101 Integrals}} — not a bad book overall, but the highlight is the appendix of 101 integrals. If you work through and understand these, I suspect you won’t struggle with a single A-level integral again. \emph{I used to do these during my regular maths lessons.}
  
\item \href{https://www.risps.co.uk/}{\textbf{RISPS}} — a collection of 40 mathematical investigations aimed at A-level students. Some are genuinely fascinating. There's also a companion set for Further Maths students. These are well worth exploring.
  

\item \href{https://www2.math.upenn.edu/~wilf/gfologyLinked2.pdf}{\textbf{Generatingfunctionology}} by Herbert Wilf — a book I discovered through the Penn course. I never finished it, but I found the parts I read both engaging and clearly written.
\end{itemize}

\subsection{MOOCs}

Many universities publish their entire courses online — and you'll likely be surprised by how accessible you find them. Over the summers of Year 11 and Year 12 (and into Year 13), I worked through the following three. I completed most of the material, and this gave me a real head start at university — I felt much more comfortable with many concepts than my peers.

\begin{enumerate}
\item \href{https://www.coursera.org/learn/single-variable-calculus}{University of Pennsylvania: Single-Variable Calculus}. This course has beautifully animated videos and offers a thorough, enriching take on A-level content. If you do only one, make it this.

\item \href{https://ocw.mit.edu/courses/18-01-calculus-i-single-variable-calculus-fall-2020/}{MIT: Single-Variable Calculus}. This overlaps heavily with the Penn course, but it's still worth doing — seeing similar content presented in different ways helps reinforce it. Plus, the exercises are different.

\item \href{https://ocw.mit.edu/courses/18-02-multivariable-calculus-fall-2007/}{MIT: Multivariable Calculus}. This is where I really learned multivariable calculus — and, perhaps tellingly, it's still the way I think about it, despite having taken the course formally at Oxford.
\end{enumerate}


\subsection{UKMT \& Olympiads}

There are many mathematical Olympiads that happen including most famously the \href{https://www.imo-official.org/}{IMO} (International Mathematical Olympiad)}. The \href{https://ukmt.org.uk/}{\textbf{UKMT}} — they run the Senior Maths Challenge and the British Mathematical Olympiad (BMO). A good target is to get a Merit in BMO Round 1 — that puts you roughly in the top 400 in your age group. Oxford and Cambridge together admit about that many maths students per year. \personal{That said, most people I knew who applied had simply achieved a high-scoring Gold in the Senior Maths Challenge (the round before the Olympiad).}

\section{Next steps}

Start by looking through the above resources and finding ones you enjoy. Maths books are best read slowly with lots of repetition. Do problems as and when you can. Try to find stuff that interests you. 

%\subsection{How our sessions can help}

%There is no longer a big need for us to cover content. The most productive use of your time with a tutor will be them explaining stuff you have already tried independently (closer to the interview also doing mock interviews). Of course going through past papers together is also useful. To facilitate this change it is probably better if I give you my whatsapp number and then you'll be able to send me stuff you get stuck on as and when it happens. I'll aim to reply as soon as possible with a solution or a useful hint.

\vspace{2cm}

\personal{While I’ve tried to present this as objectively as possible, bear in mind that it’s all based on my experience. I came 15\textsuperscript{th} in my year and I’m now going on to a PhD — but I wasn’t always the strongest student. I wasn’t even close to the top of my class until Year 13. What changed was that I started to take maths seriously — and I began working on it outside school during the summer after GCSEs. Honestly, I haven’t worked particularly hard at university. Most of what’s helped me through has been reading small bits for pleasure in the holidays. An love of the subject will get you much, much further (eventually) than any kind of natural talent or ability.}


\appendix

\section{More wider reading}\label{sec:morebooks}

\personal{I think I might have flicked through a couple of these but I don't think I seriously read any of them so can't comment on how good they are!}

\begin{itemize}

\item \href{https://isbn.is/9780297819929}{\textbf{The Magical Maze}} — Structured around the image of a maze representing interconnected mathematical ideas. Covers numbers, probability, game theory, patterns, oscillators, knots, computability, chaos, and more.

\item \href{https://isbn.is/9781846680649}{\textbf{Cabinet of Mathematical Curiosities}} — Hidden gems in logic, geometry, and probability. Includes puzzles like extracting cherries from cocktail glasses, a pop-up dodecahedron, and insights into familiar names like Pythagoras and prime numbers.

\item \href{https://isbn.is/9780747597162}{\textbf{Alex's Adventures in Numberland}} — Explores the geometry of the 50p coin, gambling strategies, mathematical patterns in nature, and the peculiar predictability of randomness.

\item \href{https://isbn.is/9781402757969}{\textbf{The Math Book}} — A historical tour through 250 short entries on major theorems and the mathematicians who discovered them. By Clifford Pickover.

\item \href{https://isbn.is/014017480X}{\textbf{You Are a Mathematician}} — Begins with the familiar and journeys through mathematical history from the Greeks to quantum theory. By David Wells.

\item \href{https://isbn.is/1862075557}{\textbf{It Must Be Beautiful}} — An accessible exploration of great equations of modern science, presented for non-specialists by various authors.

\item \href{https://isbn.is/1841155802}{\textbf{The Music of the Primes}} — A deep dive into the Riemann hypothesis and the hidden pattern of prime numbers. By Marcus du Sautoy.

\item \href{https://isbn.is/9780007214624}{\textbf{Finding Moonshine}} — A book full of insight into the nature of symmetry and the lives of those who study it. By Marcus du Sautoy.

\item \href{https://isbn.is/0691095418}{\textbf{Trigonometric Delights}} — An entertaining and historical exploration of trigonometry, blending biography and mathematics. By Eli Maor.

\item \href{https://isbn.is/9780385415804}{\textbf{The Prisoner's Dilemma}} — Explains game theory with clarity and brings the story of John von Neumann to life. By William Poundstone.

\item \href{https://isbn.is/0333780337}{\textbf{E = mc²}} — A biography of the world's most famous equation. By David Bodanis.

\item \href{https://isbn.is/9780140147391}{\textbf{Journey Through Genius}} — Explores some of the most profound theorems of mathematics and the stories behind them. By William Dunham.

\item \href{https://isbn.is/1862077355}{\textbf{The Millennium Problems}} — The seven greatest unsolved mathematical problems of our time. By Keith Devlin.

\item \href{https://isbn.is/0571212921}{\textbf{A Beautiful Mind}} — The life of Nobel laureate and mathematical genius John Nash. By Sylvia Nasar.

\item \href{https://isbn.is/0749386061}{\textbf{Chaos}} — A classic introduction to the science of chaos and unpredictability. By James Gleick.

\item \href{https://isbn.is/0140144110}{\textbf{Fractal: Images of Chaos}} — A visual exploration of fractals in computer art, population models, and planetary motion. By Hans Lauwerier.

\item \href{https://isbn.is/9780140276664}{\textbf{Six Easy Pieces}} — Richard Feynman’s clear explanation of fundamental physics concepts.

\item \href{https://isbn.is/0099736217}{\textbf{Feynman's Lost Lecture}} — A beautiful and accessible explanation of planetary motion. By David and Ruth Goodstein.

\item \href{https://isbn.is/0906212898}{\textbf{Images of Infinity}} — A creative mix of images, writings, and paradoxes related to infinity. Published by Tarquin Press.

\item \href{https://isbn.is/0883857006}{\textbf{Proofs Without Words (Book 1)}} — Visual demonstrations of mathematical truths. By Roger B. Nelson.

\item \href{https://isbn.is/0883857219}{\textbf{Proofs Without Words (Book 2)}} — A second volume of beautifully illustrated visual proofs. By Roger B. Nelson.

\item \href{https://isbn.is/1841157910}{\textbf{Fermat’s Last Theorem}} — The dramatic story of the 350-year mystery finally solved by Andrew Wiles. By Simon Singh.

\item \href{https://isbn.is/1857028899}{\textbf{The Code Book}} — The history of cryptography from the ancient world to modern-day ciphers. By Simon Singh.

\item \href{https://isbn.is/000717604X}{\textbf{The Cracking Code Book}} — A practical and engaging introduction to breaking codes. By Simon Singh.

\item \href{https://isbn.is/0140256024}{\textbf{Does God Play Dice?}} — An excellent introduction to chaos theory. The title quotes Einstein’s skepticism. By Ian Stewart.

\item \href{https://isbn.is/0140132376}{\textbf{Game, Set and Math}} — Mathematical insights into tennis, puzzles, and beyond. By Ian Stewart.

\item \href{https://isbn.is/0007214227}{\textbf{Longitude}} — The true story of the quest to solve the longitude problem at sea. By Dava Sobel.

\item \textbf{Mathematical Puzzles and Diversions}, \textbf{More Mathematical Puzzles and Diversions}, \textbf{Mathematical Circus}, \textbf{The Unexpected Hanging}, \textbf{Fractal Music, Hypercards and More} — All by Martin Gardner. A treasure trove of mathematical amusement and insight.

\item \href{http://www.geom.uiuc.edu/~banchoff/Flatland/}{\textbf{Flatland – A Romance of Many Dimensions}} — A story of a square living in a two-dimensional world and his encounter with a sphere. By Edwin A. Abbott. Also available to read online.

\end{itemize}




\end{document}

